---
title: "CARTEpigenoQC Report"
output: html_document
params:
  sample_name: "PBMC_10k"
  frip_file: "../../data/example_data/frip_results.tsv"
  insert_file: "../../data/example_data/car_coverage.tsv"
---

```{r setup-global-unique, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(data.table)
library(ggplot2)
library(DT)
library(gridExtra)
```

# 🧬 Sample Information

- **Sample name**: `r params$sample_name`
- **Report date**: `r Sys.Date()`

---

# 📊 FRiP (Fraction of Reads in Peaks)

```{r frip-load-unique}
frip_dt <- fread(params$frip_file)
if (!dir.exists("figures")) dir.create("figures", recursive = TRUE)
```

## FRiP Summary Statistics

```{r frip-summary-unique}
summary(frip_dt$frip)
```

## FRiP Distribution Plot

```{r frip-plot-unique}
frip_plot <- ggplot(frip_dt, aes(x = frip)) +
  geom_histogram(fill = "steelblue", bins = 50) +
  labs(title = "FRiP Score Distribution", x = "FRiP Score", y = "Number of Cells") +
  theme_minimal()
ggsave("../../figures/example_frip_plot.png", frip_plot, width = 6, height = 4, bg = "white")
frip_plot
```

## Top Cells by FRiP

```{r frip-top-table-unique}
top_frip <- frip_dt[order(-frip)][1:20]
datatable(top_frip, options = list(pageLength = 10))
png("../../figures/frip_top_table.png", width = 800, height = 400)
grid.table(top_frip)
dev.off()
```

---

# 📌 CAR Insertion Site Coverage

```{r insertions-load-unique}
insert_dt <- fread(params$insert_file)
```

## Insertion Site Coverage Heatmap

```{r insertions-heatmap-unique}
insert_heatmap <- ggplot(insert_dt, aes(x = site, y = barcode, fill = insert_reads)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "firebrick") +
  labs(title = "Per-cell CAR Insertion Site Coverage", x = "Insertion Site", y = "Cell Barcode") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
ggsave("../../figures/car_coverage_heatmap.png", insert_heatmap, width = 7, height = 5, bg = "white")
insert_heatmap
```

## Top Sites by Total Read Count

```{r insertions-top-sites-unique}
top_sites <- insert_dt[, .(total_reads = sum(insert_reads)), by = site][order(-total_reads)][1:10]
print(top_sites)
png("../../figures/car_site_table.png", width = 800, height = 400)
grid.table(top_sites)
dev.off()
```

---

# ✅ Summary

- Total cells analyzed: `r nrow(frip_dt)`
- Insertion sites detected: `r length(unique(insert_dt$site))`
- Report generated on: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`
